---
title: "World Happiness"
subtitle: "2022"
author: "Ben Mitchell, Eddie Jiang, Callum Knox, Sarah Brown-Scott"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(tidymodels)
library(broom)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
style_mono_dark(base_color = "ivory")
```

```{r load-data, include=FALSE}
happiness <- read_csv("presentation_files/2015.csv")
countries <- read_csv("presentation_files/countries of the world.csv")
```

```{r include=FALSE}
countries$Country[countries$Country == "Cote d\'Ivoire"] = "Ivory Coast"
countries$Country[countries$Country == "Bosnia & Herzegovina"] = "Bosnia and Herzegovina"
countries$Country[countries$Country == "Central African Rep."] = "Central African Republic"
countries$Country[countries$Country == "Congo, Dem. Rep."] = "Congo (Kinshasa)"
countries$Country[countries$Country == "Congo, Repub. of the"] = "Congo (Brazzaville)"
```

```{r include=FALSE}
happiness_countries <- left_join(happiness, countries, by = "Country")
```

```{r include=FALSE}
colnames(happiness_countries)[4] <- "happiness_score"
colnames(happiness_countries)[6] <- "economy"
colnames(happiness_countries)[7] <- "family"
colnames(happiness_countries)[8] <- "health"
colnames(happiness_countries)[10] <- "trust"
colnames(happiness_countries)[9] <- "freedom"
colnames(happiness_countries)[20] <- "gdp"
colnames(happiness_countries)[19] <- "infant_mortality"
colnames(happiness_countries)[22] <- "phones"
```



class: centre, middle

##Question & Hypothesis
---

class: left, middle
#Question

-The research question we will be investigating is, 'How  Do a Range of Factors Affect Worldwide Happiness?'.

- We selected this question as mental health is a prominent issue in society and analysis of the data in this way could give some indicators of ways to improve the overall happiness of a country.

---

class: left, middle

- We hypothesise that in countries with higher GDP and freedom rating , the happiness score will be higher than countries with a lower GDP and freedom rating.

-We hypothesise that in countries where Government Corruption is higher, the happiness score will be lower than in countries with less government corruption.

---

class: left, middle

# Dataset Information

-To try and answer this question we picked 2 data sets.

-They were "World happiness Report" from 2015, and "Countries of The World.

-World Happiness Report contains data from a survey about overall happiness in 158 countries

```{r echo=FALSE}
glimpse(happiness)
```

- Variables economy, family, health, freedom, trust, and generosity show how happy people are with these things.

---

class: left, middle

-Countries of The World has general statistics about all countires.

-GDP, literacy%, population density etc.

-We combined the two data sets as we thought that this would help answer the question.

-Combined sets has opinions and facts to compare.
---

class: left, middle

#Countries

```{r echo=FALSE}
glimpse(countries)
```
---
class: left, middle

#Summary Statistics

Variable | Mean | Max Value | Min Value | Standard Deviation
--------|---------|---------|-----------|---------------------
`Happiness Score` | `5.38` |  `7.59`  | `2.84`  | `1.145`
`Health` | `0.63` | `1.03`  | `0`  | `0.247`
`Economy` | `0.85` |  `1.69` |  `0` | `0.403`
`Trust` | `0.143`  |  `0.552` | `0`  | `0.120`
`Family` | `0.991`  | `1.4`  |  `0` | `0.272`
`Freedom` |  `0.429` | `0.668`  | `0`  | `0.151`

---

class: left, center

##Plots

---
class: left, center

-As Family Score increases Happiness Score also Increases

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = family, y = happiness_score))+
  geom_jitter() +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Family",
       x = "Family",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```


---

class: left, center

-As Freedom Score increases Happiness Score also Increases

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = freedom, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Freedom",
       x = "Freedom",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```


---
class: left, center

-As Health Score increases Happiness Score also Increases

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = health, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Health",
       x = "Health",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```


---
class: left, center

-As Trust Score increases Happiness Score also Increases

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = trust, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Government Trust",
       x = "Trust",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```


---
class: left, center

-As Economy Score increases Happiness Score also Increases

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = economy, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Economy", 
       x = "Economy",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```
---
class: left, center

-Some regions are happier on average than others

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness_countries %>%
  ggplot(data = happiness_countries, 
         mapping = aes(x = happiness_score, y = Region.x))+
  geom_boxplot()+
  labs(title = "Happiness Score shown by Region",
       subtitle = "Regions are split into sub regions",
       x = "Happiness Score",
       y = "Region") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        plot.subtitle = element_text(size = rel(1),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```
---
class: left, center

##Models

---
class: left, center

-We are 95% confident that for each additional  0.1 of a point of economy score, we would expect gdp to increase by $1734.00 to $2390.46, on average

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness %>%
  ggplot(data = happiness_countries, 
         mapping = aes(x = economy, y = gdp))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Economy Score against GDP",
       x = "Economy Score",
       y = "GDP") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```
---
class: left, center

-we are 95% confident that for each additional  0.1 of a point of health score, we would expect infant mortality to decrease by 963.71 to 1418.07, on average.

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness_countries %>%
    drop_na(infant_mortality) %>%
  ggplot(data = happiness_countries,
         mapping = aes(x = health, y = infant_mortality,
                       na.rm = TRUE))+
  geom_jitter() +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Infant Mortality against Health",
       x = "Health Score",
       y = "Infant Mortality") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```
---
class: left, center

-No correlation.

```{r echo=FALSE, message=FALSE, warning=FALSE}
happiness_countries %>%
  drop_na(phones) %>%
  ggplot(data = happiness_countries,
         mapping = aes(x = phones, y = happiness_score,
                       na.rm = TRUE))+
  geom_jitter() +
  expand_limits(y = c(0, 10)) +
  labs(title = "Happiness Score against Phones per 1000 People",
       x = "Phones Per 1000 People",
       y = "Happiness Score") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```
---
class: middle, center

##Conclusion & Limitations

---
class: left, middle

#Limitations

-People surveyed.

-Issues in the country may affect results of survey.

-Outdated
---
class: left, middle

#Conclusions

-"We hypothesise that in countries with higher GDP and freedom rating , the happiness score will be higher than countries with a lower GDP and freedom rating."

-"We hypothesise that in countries where Government Corruption is higher, the happiness score will be lower than in countries with less government corruption."

-Most graphs showed expected trends.

-Certain factors could be seen to affect overall happiness.

-Some factors not in the dataset will affect happiness

---



```{r include=FALSE, eval = FALSE}
#model predicting economy score from GDP
set.seed(1234)

economy_boot <- bootstraps(happiness_countries, times = 5000)

economy_models <- economy_boot %>%
  mutate(
    model = map(splits, ~lm(gdp ~ economy, data = .)),
    coef_info = map(model, tidy)
  )

economy_coef <- economy_models %>%
  unnest(coef_info)

economy_confidence <- int_pctl(economy_models, coef_info)
```

```{r include=FALSE, eval = FALSE}
#model predicting health score from infant mortality
set.seed(1234)

health_boot <- bootstraps(happiness_countries, times = 5000)

health_models <- health_boot %>%
  mutate(
    model = map(splits, ~lm(infant_mortality ~ health, data = .)),
    coef_info = map(model, tidy)
  )

haelth_coef <- health_models %>%
  unnest(coef_info)

health_confidence <- int_pctl(health_models, coef_info)
```

```{r include=FALSE, eval = FALSE}
#model predicting happiness from phones per 1000 people
set.seed(7399)

phones_boot <- bootstraps(happiness_countries, times = 5000)

phones_models <- phones_boot %>%
  mutate(
    model = map(splits, ~lm(happiness_score ~ phones, data = .)),
    coef_info = map(model, tidy)
  )

phones_coef <- phones_models %>%
  unnest(coef_info)

phones_confidence <- int_pctl(phones_models, coef_info)
```

```{r include=FALSE, eval = FALSE}
phones_confidence_mean <- phones_confidence %>%
  filter(term != "(Intercept)")

lower <- mean(phones_confidence_0$.lower)
upper <- mean(phones_confidence_0$.upper)
```